---
title: "Untitled"
author: 'xxxx'
date: "2023-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
set.seed(1)
```

## 1. Helper functions for basic simulation

### a) Homogeneous Poisson processes (join, accident)

```{r}
# this function simulates events and their timing, according to a homogeneous poisson process, and outputs the number of events that occur within the time horizon specified
num_hom = function(horizon, rate) {
  
  # initialise
  t = 0 # current time
  I = 0 # number of events
  S = c() # will hold event times
  
  # simulate events and stop when current event time surpasses the time horizon
  while (t <= horizon){
    
    # generate time of next event
    U = runif(1)
    t = t - (1/rate) * log(U)
    
    # only increment count of events if new event time does not surpass the time horizon
    if (t <= horizon) {
      I <- I + 1
      S <- c(S, t)
    }
  }
  
  # return the number of events that occured
  return(I)
}
```

### b) Non-homogeneous Poisson process (leave)

```{r}
# this function simulates events and their timing, according to a non-homogeneous poisson process, and outputs the number of events that occur within the time horizon specified
num_nonhom_leave = function(horizon) {
  
  # initialise
  t = 0 # current time
  I = 0 # number of events
  S = c() # will hold event times
  
  mu_bound <- 1/12 # upper bound for non-homogeneous poisson process with rate mu(t)

  # simulate events and stop when current event time surpasses the time horizon
  while (t <= horizon){
    
    # generate time of next event
    U = runif(1)
    t = t - (1/mu_bound) * log(U)
    
    # calculate likelihood of an event occurring at the current time
    L <- mu(t) / mu_bound

    # only increment the count of events if the generated random variable is less than or equal to the likelihood
    if (runif(1) <= L){
      I = I + 1
      S = c(S,t)
    } 
  }
  
  # return the number of events that occured
  return(I)
}
```

### c) Bernoulli r.v. (if claim)

```{r}
# this function simulates whether or not a claim is made, based on a Bernoulli distributed random variable, with probability of success (claim made) at 0.6
rv_bernoulli <- function(p_success) {
  return(runif(1) < p_success)
}
```

### d) Specific discrete r.v. (amount claimed)

```{r}
# this function simulates the amount claimed on an accident, based on the given function 300X+500, where X takes values {0, 1, ..., 10} and follows the uniform distribution 
rv_specific_discrete <- function() {
  X <- floor(11 * runif(1))
  return(300 * X + 500)
}
```

## 2. Initialisation of parameters

```{r}
n_0 <- 0 # initial number of customers
c_0 <- 50000 # initial capital stock
M <- 300 # fixed monthly fee
C <- 30000 # capital threshold before legal complications
lambda <- 3 # monthly rate at which customers join 
mu <- function(t) 1/(12 + t) # monthly rate at which customers leave 
alpha <- 3.5 # annual rate of accidents
K <- 1000 # the number of simulation iterations
```

## 3. Main code for simulation (likelihood)

note: must reset monthly, due to the membership fee (M) being paid monthly

### a) Simulation

```{r}
# initialize results and counters

verdict <- c() # stores 0s and 1s where a 1 indicates company 'failed'
capital <- c() # stores the capital available at the end of the 12 months

for (i in 1:K) {
  
  # initialise 
  t <- 0 # current time
  n <- n_0 # starting number of customers
  c <- c_0 # starting amount of capital

  # repeat for each of the 12 months in a year
  for (month in 1:12) {
    
    # ARRIVALS (per month)
    num_join_month = num_hom(1, lambda)
    
    # DEPARTURES (per month)
    num_leave_month = num_nonhom_leave(1)
    
    # -update- capital (based on number of clients and their membership payment (per month))
    n <- n + num_join_month - num_leave_month
    c <- c + n*M
    
    # ACCIDENTS (per customer per month)
    num_acc_month <- 0
    for (j in 1:n) {
      num_acc_customer <- num_hom(1, alpha/12)
      
      # -update- accidents counter
      num_acc_month <- num_acc_month + num_acc_customer
    }

    # CLAIM & AMOUNT (per accident)
    for (k in 1:num_acc_month) {
      
      # claim, per accident
      claim_made_customer = rv_bernoulli(0.6)
      
      # amount, per accident
      if (claim_made_customer) {
        amount_claimed_customer = rv_specific_discrete()
        
        # -update- capital (based on each accident claimed)
        c <- c - amount_claimed_customer
      }
    }
  }
  
  # -update- verdict (for current simulation year) 
  if (c < C) {
    verdict = c(verdict, 1) # record 'company failed' as 1
  } else {
    verdict = c(verdict, 0) # record 'company succeeded' as 0
  }
  # -update- capital at the end of the year
  capital <- c(capital, c)
}
```

### b) [Likelihood] Calculations and results

```{r}
# probability of 'failure' (mean)
likelihood = sum(verdict) / K
paste0("Estimated likelihood of legal complications (%): ", likelihood*100)
```
